---
title: "Composing Tracks"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 6,
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(ezGenomeTracks)
library(ggplot2)
```

One of the key features of ezGenomeTracks is the ability to compose multiple tracks into a single, publication-ready figure. This vignette covers vertical and horizontal stacking, track height control, and adding annotations.

# Vertical Stacking with vstack_plot()

The most common composition is stacking tracks vertically with a shared genomic x-axis.

## Basic Stacking

```{r basic-stack}
# Define region
region <- "chr2:231109000-231115000"

# Load data
bw_file <- system.file(
  "extdata",
  "avg_chr2-231091223_231109786_231113600_0.bw",
  package = "ezGenomeTracks"
)

# Create individual tracks
coverage_track <- ez_coverage(
  bw_file,
  region = region,
  colors = "steelblue",
  y_axis_style = "simple"
)

# Simulated peaks for this region
peaks_df <- data.frame(
  seqnames = "chr2",
  start = c(231109500, 231111000, 231113500),
  end = c(231110500, 231112500, 231114500),
  score = c(50, 100, 75)
)

peak_track <- ez_feature(
  peaks_df,
  region = region,
  use_score = TRUE,
  fill = "darkred"
)

# Stack them
vstack_plot(
  coverage_track,
  peak_track,
  region = region
)
```

## Controlling Track Heights

Use the `heights` parameter to control relative track heights. Heights are specified for tracks 2 onward (relative to track 1).

```{r heights}
# More emphasis on coverage, less on peaks
vstack_plot(
  coverage_track,
  peak_track,
  region = region,
  heights = c(0.3)  # peak track is 0.3x the height of coverage
)
```

## Stacking Multiple Tracks

```{r multi-stack, fig.height=8}
# Load all three genotype bigWigs
bw_files <- list(
  "AA" = system.file("extdata", "avg_chr2-231091223_231109786_231113600_0.bw", package = "ezGenomeTracks"),
  "AG" = system.file("extdata", "avg_chr2-231091223_231109786_231113600_1.bw", package = "ezGenomeTracks"),
  "GG" = system.file("extdata", "avg_chr2-231091223_231109786_231113600_2.bw", package = "ezGenomeTracks")
)

# Create a multi-track coverage plot
multi_coverage <- ez_coverage(
  bw_files,
  region = region,
  colors = c("blue2", "orange2", "purple2"),
  y_axis_style = "minmax",
  facet_label_position = "left"
)

# Add peaks and interactions
interactions_df <- data.frame(
  start1 = c(231109800, 231111500),
  end1 = c(231110200, 231112000),
  start2 = c(231113000, 231114000),
  end2 = c(231113500, 231114500),
  score = c(20, 35)
)

arc_track <- ez_link(
  interactions_df,
  region = region,
  direction = "down",
  colors = "gray40"
)

# Stack all
vstack_plot(
  multi_coverage,
  arc_track,
  peak_track,
  region = region,
  heights = c(0.5, 0.3)
)
```

# Horizontal Stacking with hstack_plot()

Use `hstack_plot()` to place plots side by side, useful for comparing different genomic regions.

## Comparing Regions

```{r hstack, fig.height=3}
# Two different regions
region1 <- "chr2:231109000-231112000"
region2 <- "chr2:231112000-231115000"

plot1 <- ez_coverage(
  bw_files[[1]],
  region = region1,
  colors = "steelblue",
  y_axis_style = "simple"
)

plot2 <- ez_coverage(
  bw_files[[1]],
  region = region2,
  colors = "steelblue",
  y_axis_style = "simple"
)

# Side by side
hstack_plot(plot1, plot2)
```

## Controlling Widths

```{r hstack-widths, fig.height=3}
# Make left plot wider
hstack_plot(plot1, plot2, widths = c(2))  # Right plot is 2x width of left
```

# Adding Annotations

ezGenomeTracks provides helper functions to add annotations to any track.

## Vertical Lines

Highlight specific genomic positions (e.g., SNP locations, TSS).

```{r vline}
# Add a vertical line to mark a position
annotated_track <- add_vline(
  coverage_track,
  position = 231111500,
  color = "red",
  linetype = "dashed"
)

annotated_track
```

## Horizontal Lines

Mark thresholds or reference values.

```{r hline}
# Add a threshold line
annotated_track <- add_hline(
  coverage_track,
  y = 50,
  color = "red",
  linetype = "dotted"
)

annotated_track
```

## Highlight Regions

Draw rectangles to highlight genomic regions of interest.

```{r rect}
# Highlight a region
annotated_track <- add_rect(
  coverage_track,
  xmin = 231110500,
  xmax = 231112000,
  fill = "yellow",
  alpha = 0.3
)

annotated_track
```
## Text Labels

Add text annotations at specific positions.

```{r text}
# Add a label
annotated_track <- add_text(
  coverage_track,
  x = 231111000,
  y = 60,
  label = "Peak",
  color = "red",
  size = 4
)

annotated_track
```

## Combining Multiple Annotations

```{r multi-annot}
# Chain multiple annotations
annotated_track <- coverage_track |>
  add_rect(xmin = 231110500, xmax = 231112000, fill = "yellow", alpha = 0.2) |>
  add_vline(position = 231111000, color = "red") |>
  add_text(x = 231111000, y = 70, label = "SNP", color = "red", size = 3)

annotated_track
```

# Real-World Examples

## QTL Visualization

Combine Manhattan plot with coverage tracks to show genetic associations in context.

```{r qtl-example, fig.height=7}
region <- "chr2:231109000-231115000"

# Simulated GWAS data for this region
gwas_df <- data.frame(
  CHR = 2,
  BP = seq(231109000, 231115000, by = 200),
  P = 10^(-runif(31, 0, 4)),
  SNP = paste0("rs", 1:31)
)
gwas_df$P[15] <- 1e-6  # Add a significant SNP

# Manhattan track
manhattan_track <- ez_manhattan(
  gwas_df,
  region = region,
  chr = "CHR",
  bp = "BP",
  p = "P",
  y_axis_style = "full",
  threshold_p = 5e-5,
  threshold_color = "red"
)

# Coverage tracks
coverage_multi <- ez_coverage(
  bw_files,
  region = region,
  colors = c("blue2", "orange2", "purple2"),
  y_axis_style = "minmax",
  facet_label_position = "left"
)

# Stack
vstack_plot(
  manhattan_track,
  coverage_multi,
  region = region,
  heights = c(1.5)
)
```

## Sashimi Plot with Gene Track

```{r sashimi-example, fig.height=6}
region <- "chr2:231107879-231115507"

# Load sashimi data
bw0 <- system.file("extdata", "avg_chr2-231091223_231109786_231113600_0.bw", package = "ezGenomeTracks")

link0 <- data.table::fread(
  system.file("extdata", "chr2-231091223_231109786_231113600_0.sashimi", package = "ezGenomeTracks"),
  header = FALSE,
  col.names = c("seqnames", "start", "end", "name", "score", "strand")
)

# Sashimi track
sashimi_track <- ez_sashimi(
  coverage_data = bw0,
  junction_data = link0,
  region = region,
  colors = "purple2",
  linewidth_range = c(0.1, 0.5),
  height_factor = 0.03,
  junction_curvature = 0.03
)

# Display
sashimi_track
```

# Tips and Best Practices

1. **Track order**: Place tracks that need more vertical space (Hi-C, coverage) at the top, compact tracks (features, genes) at the bottom.

2. **Consistent regions**: Always use the same region string for all tracks in a stack to ensure proper x-axis alignment.

3. **Heights**: Start with equal heights and adjust iteratively. Coverage tracks often need 2-3x the height of feature/gene tracks.

4. **Annotations**: Add annotations to individual tracks before stacking for proper positioning.

5. **Figure dimensions**: Set appropriate `fig.height` and `fig.width` in your R Markdown chunks or `ggsave()` dimensions for publication.

6. **Legend placement**: Use `show_legend = TRUE` selectively on one track to avoid redundant legends.
