---
title: "Showcase: Multi-omic Visualization with plotgardenerData"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 6,
  message = FALSE,
  warning = FALSE
)

# Check if plotgardenerData is available
has_plotgardener <- requireNamespace("plotgardenerData", quietly = TRUE) &&
  requireNamespace("TxDb.Hsapiens.UCSC.hg19.knownGene", quietly = TRUE)
```

This vignette demonstrates ezGenomeTracks using realistic, publication-quality data from the [plotgardenerData](https://bioconductor.org/packages/plotgardenerData/) Bioconductor package. These examples showcase how to create comprehensive multi-omic genome browser views combining Hi-C, ChIP-seq, chromatin loops, and GWAS data.

# Prerequisites

First, install the required packages from Bioconductor:

```{r install, eval=FALSE}
if (!requireNamespace("BiocManager", quietly = TRUE))
    install.packages("BiocManager")

BiocManager::install("plotgardenerData")
BiocManager::install("TxDb.Hsapiens.UCSC.hg19.knownGene")
```

```{r setup}
library(ezGenomeTracks)
library(ggplot2)
```

```{r load-data, eval=has_plotgardener}
library(plotgardenerData)
library(TxDb.Hsapiens.UCSC.hg19.knownGene)

# Load the datasets we'll use
data("IMR90_HiC_10kb")
data("GM12878_HiC_10kb")
data("IMR90_DNAloops_pairs")
data("IMR90_ChIP_H3K27ac_signal")
data("GM12878_ChIP_H3K27ac_signal")
data("IMR90_ChIP_CTCF_signal")
data("hg19_insulin_GWAS")

# Set up TxDb for gene annotation
txdb <- TxDb.Hsapiens.UCSC.hg19.knownGene

# --- Preprocess data to match ezGenomeTracks expected formats ---

# Hi-C data: plotgardenerData uses duplicate column names, rename to pos1/pos2/score
preprocess_hic <- function(hic_df) {
  colnames(hic_df) <- c("pos1", "pos2", "score")
  hic_df
}
IMR90_HiC_10kb <- preprocess_hic(IMR90_HiC_10kb)
GM12878_HiC_10kb <- preprocess_hic(GM12878_HiC_10kb)

# Signal data: rename 'chrom' to 'seqnames' for consistency
preprocess_signal <- function(signal_df) {
  if ("chrom" %in% colnames(signal_df)) {
    colnames(signal_df)[colnames(signal_df) == "chrom"] <- "seqnames"
  }
  signal_df
}
IMR90_ChIP_H3K27ac_signal <- preprocess_signal(IMR90_ChIP_H3K27ac_signal)
GM12878_ChIP_H3K27ac_signal <- preprocess_signal(GM12878_ChIP_H3K27ac_signal)
IMR90_ChIP_CTCF_signal <- preprocess_signal(IMR90_ChIP_CTCF_signal)

# GWAS data: rename columns for ez_manhattan
colnames(hg19_insulin_GWAS) <- c("chr", "bp", "p", "snp", "LD")
# Remove 'chr' prefix for consistency
hg19_insulin_GWAS$chr <- gsub("^chr", "", hg19_insulin_GWAS$chr)
```

# Available Datasets

plotgardenerData provides several datasets from ENCODE cell lines:
| Dataset | Type | Cell Line | Description |
|---------|------|-----------|-------------|
| `IMR90_HiC_10kb` | Hi-C | IMR90 | 10kb resolution contact matrix |
| `GM12878_HiC_10kb` | Hi-C | GM12878 | 10kb resolution contact matrix |
| `IMR90_DNAloops_pairs` | Loops | IMR90 | Chromatin loop calls |
| `IMR90_ChIP_H3K27ac_signal` | Signal | IMR90 | H3K27ac ChIP-seq |
| `GM12878_ChIP_H3K27ac_signal` | Signal | GM12878 | H3K27ac ChIP-seq |
| `IMR90_ChIP_CTCF_signal` | Signal | IMR90 | CTCF ChIP-seq |
| `hg19_insulin_GWAS` | GWAS | - | Insulin-related associations |

All data is aligned to **hg19** and covers regions on **chromosome 21**.

# Example 1: Hi-C Triangle with Signal Track

The most common multi-omic view combines Hi-C contact data with ChIP-seq signal.

```{r hic-signal, fig.height=7, eval=has_plotgardener}
region <- "chr21:28000000-30000000"

# Hi-C track (triangle view)
hic_track <- ez_hic(
  IMR90_HiC_10kb,
  region = region,
  resolution = 10000,
  style = "triangle",
  palette = "cooler",
  transform = "log10"
)

# H3K27ac signal track
h3k27ac_track <- ez_coverage(
  IMR90_ChIP_H3K27ac_signal,
  region = region,
  colors = "orange2",
  y_axis_style = "simple"
)

# Gene track
gene_track <- ez_gene(txdb, region = region)

# Stack all tracks
vstack_plot(
  hic_track,
  h3k27ac_track,
  gene_track,
  region = region,
  heights = c(1, 0.5)
)
```

# Example 2: Comparing Cell Types

Compare chromatin architecture between IMR90 (fibroblast) and GM12878 (lymphoblastoid) cells.

```{r compare-cells, fig.height=8, eval=has_plotgardener}
region <- "chr21:28500000-29500000"

# IMR90 tracks
hic_imr90 <- ez_hic(
  IMR90_HiC_10kb,
  region = region,
  resolution = 10000,
  style = "triangle",
  palette = "cooler",
  transform = "log10",
  max_distance = 500000
)

signal_imr90 <- ez_coverage(
  IMR90_ChIP_H3K27ac_signal,
  region = region,
  colors = "darkorange",
  y_axis_style = "minmax"
)

# GM12878 tracks
hic_gm12878 <- ez_hic(
  GM12878_HiC_10kb,
  region = region,
  resolution = 10000,
  style = "triangle",
  palette = "cooler",
  transform = "log10",
  max_distance = 500000
)

signal_gm12878 <- ez_coverage(
  GM12878_ChIP_H3K27ac_signal,
  region = region,
  colors = "purple3",
  y_axis_style = "minmax"
)

# Gene track (shared)
genes <- ez_gene(txdb, region = region)

# Create IMR90 view
imr90_view <- vstack_plot(
  hic_imr90,
  signal_imr90,
  region = region,
  heights = c(0.5)
)

# Create GM12878 view
gm12878_view <- vstack_plot(
  hic_gm12878,
  signal_gm12878,
  region = region,
  heights = c(0.5)
)

# Display IMR90 view
imr90_view
```

```{r compare-cells-gm, fig.height=5, eval=has_plotgardener}
# Display GM12878 view
gm12878_view
```

# Example 3: Chromatin Loops

Chromatin loops connect distant regulatory elements. The `ez_link()` function visualizes these interactions as arcs.

```{r loops, fig.height=3, eval=FALSE}
region <- "chr21:28000000-30000000"

# Chromatin loop track showing DNA interactions
ez_link(
  IMR90_DNAloops_pairs,
  region = region,
  direction = "down",
  colors = "blue3",
  size = 0.8,
  alpha = 0.7
)
```

*Note: This example is shown as code only due to a rendering limitation with bezier curves in knitr documents. Run this code interactively to see the loop visualization.*

The loops connect enhancer-promoter pairs and indicate regions of 3D chromatin contact.

# Example 3b: Multi-track Epigenomic View

Combine Hi-C contact frequency with ChIP-seq signal tracks to interpret chromatin architecture.

```{r multitrack, fig.height=8, eval=has_plotgardener}
region <- "chr21:28000000-30000000"

# Hi-C track
hic_track <- ez_hic(
  IMR90_HiC_10kb,
  region = region,
  resolution = 10000,
  style = "triangle",
  palette = "cooler",
  transform = "log10",
  max_distance = 1000000
)

# H3K27ac signal (active enhancers/promoters)
h3k27ac_track <- ez_coverage(
  IMR90_ChIP_H3K27ac_signal,
  region = region,
  colors = "orange2",
  y_axis_style = "simple"
)

# CTCF signal (architectural protein marking loop anchors)
ctcf_track <- ez_coverage(
  IMR90_ChIP_CTCF_signal,
  region = region,
  colors = "green4",
  y_axis_style = "simple"
)

# Gene track
gene_track <- ez_gene(txdb, region = region)

# Stack all tracks
vstack_plot(
  hic_track,
  h3k27ac_track,
  ctcf_track,
  gene_track,
  region = region,
  heights = c(0.5, 0.5, 0.5)
)
```

# Example 4: GWAS Integration

Combine genetic association data with epigenomic context to interpret GWAS signals.

```{r gwas, fig.height=7, eval=has_plotgardener}
# Filter GWAS data to a region with signal
gwas_region <- "chr21:28500000-29500000"

# Regional Manhattan plot
manhattan_track <- ez_manhattan(
  hg19_insulin_GWAS,
  region = gwas_region,
  y_axis_style = "full",
  size = 1.5,
  threshold_p = 5e-6,
  threshold_color = "red"
)

# H3K27ac track - active enhancers may harbor causal variants
h3k27ac_track <- ez_coverage(
  IMR90_ChIP_H3K27ac_signal,
  region = gwas_region,
  colors = "orange2",
  y_axis_style = "simple"
)

# CTCF track - may indicate regulatory boundaries
ctcf_track <- ez_coverage(
  IMR90_ChIP_CTCF_signal,
  region = gwas_region,
  colors = "green4",
  y_axis_style = "simple"
)

# Gene track
gene_track <- ez_gene(txdb, region = gwas_region)

# Create integrated QTL view
vstack_plot(
  manhattan_track,
  h3k27ac_track,
  ctcf_track,
  gene_track,
  region = gwas_region,
  heights = c(0.8, 0.8, 0.5)
)
```

# Example 5: Complete Multi-omic Browser

A comprehensive view combining Hi-C, ChIP-seq, and gene annotations for publication-quality figures.

```{r complete, fig.height=9, eval=has_plotgardener}
region <- "chr21:28500000-29500000"

# Hi-C (chromatin architecture)
p_hic <- ez_hic(
  IMR90_HiC_10kb,
  region = region,
  resolution = 10000,
  style = "triangle",
  max_distance = 500000,
  palette = "cooler",
  transform = "log10"
)

# H3K27ac (active enhancers/promoters)
p_h3k27ac <- ez_coverage(
  IMR90_ChIP_H3K27ac_signal,
  region = region,
  colors = "orange2",
  y_axis_style = "minmax"
)

# CTCF (insulator/architectural)
p_ctcf <- ez_coverage(
  IMR90_ChIP_CTCF_signal,
  region = region,
  colors = "green4",
  y_axis_style = "minmax"
)

# Genes
p_genes <- ez_gene(txdb, region = region)

# Final composition
vstack_plot(
  p_hic,
  p_h3k27ac,
  p_ctcf,
  p_genes,
  region = region,
  heights = c(0.4, 0.4, 0.5)
)
```

# Tips for Working with plotgardenerData

1. **Assembly**: All plotgardenerData is aligned to **hg19**. Ensure your TxDb and other annotations match.

2. **Resolution**: Hi-C data is at 10kb resolution. Adjust your region size accordingly (at least 500kb-2Mb for meaningful patterns).

3. **Memory**: Hi-C matrices can be large. Use `max_distance` to limit data when visualizing narrow regions.

4. **Recommended regions**: The data has best coverage around chr21:28,000,000-30,300,000.

5. **Cell type context**: IMR90 are lung fibroblasts; GM12878 are lymphoblastoid cells. Differences reflect cell type-specific chromatin organization.

# Data Not Available

```{r no-data, eval=!has_plotgardener, echo=FALSE, results='asis'}
cat("
**Note**: This vignette requires the `plotgardenerData` package from Bioconductor.

To run these examples, install the required packages:
```r
BiocManager::install('plotgardenerData')
BiocManager::install('TxDb.Hsapiens.UCSC.hg19.knownGene')
```

Then re-render this vignette.
")
```
