---
title: "tracks_sashimi"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ezGenomeTracks)
```

Sashimi plots are an elegant way to visualize mRNA splicing data, but few existing tools can do it. Years ago I wrote a custom track class for pyGenomeTracks and have seen several people using it. Here, `ezGenomeTracks` provides a unified `ez_sashimi` function that makes it even easier to plot Sashimi plots. It takes a bigwig file and a link file as inputs.

```{r}
bw0 <- system.file(
  "extdata",
  "avg_chr2-231091223_231109786_231113600_0.bw",
  package = "ezGenomeTracks"
)

bw1 <- system.file(
  "extdata",
  "avg_chr2-231091223_231109786_231113600_1.bw",
  package = "ezGenomeTracks"
)

bw2 <- system.file(
  "extdata",
  "avg_chr2-231091223_231109786_231113600_2.bw",
  package = "ezGenomeTracks"
)

link0 <- data.table::fread(
  system.file(
    "extdata",
    "chr2-231091223_231109786_231113600_0.sashimi",
    package = "ezGenomeTracks"
  ),
  header = F,
  col.names = c("seqnames", "start", "end", "name", "score", "strand")
)

link1 <- data.table::fread(
  system.file(
    "extdata",
    "chr2-231091223_231109786_231113600_1.sashimi",
    package = "ezGenomeTracks"
  ),
  header = F,
  col.names = c("seqnames", "start", "end", "name", "score", "strand")
)

link2 <- data.table::fread(
  system.file(
    "extdata",
    "chr2-231091223_231109786_231113600_2.sashimi",
    package = "ezGenomeTracks"
  ),
  header = F,
  col.names = c("seqnames", "start", "end", "name", "score", "strand")
)
```

```{r}
head(link0)
```

```{r, fig.height=1.5}
ez_sashimi(
  coverage_data = bw0,
  junction_data = link0 |> dplyr::mutate(score = signif(score, 3)),
  linewidth_range = c(0.1, 0.5),
  height_factor = 0.03,
  region = "chr2:231107879-231115507",
  junction_curvature = 0.03
)
```

We can plot multiple track together as we do for `ez_coverage`.

```{r}
ez_sashimi(
  coverage_data = list(bw0, bw1, bw2),
  junction_data = list(
    link0 |> dplyr::mutate(score = signif(score, 3)),
    link1 |> dplyr::mutate(score = signif(score, 3)),
    link2 |> dplyr::mutate(score = signif(score, 3))
  ),
  linewidth_range = c(0.1, 0.5),
  height_factor = 0.03,
  region = "chr2:231107879-231115507",
  junction_curvature = 0.03,
  colors = c("purple2", "orange2", "blue2")
)
```