---
title: "Manhattan plot"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ezGenomeTracks)
```

`ezGenomeTracks` provides two specialized functions for visualizing GWAS association data:

- **`ez_manhattan()`** - Genome-wide Manhattan plots across multiple chromosomes
- **`ez_locusZoom()`** - Regional association plots (LocusZoom-style) for a single locus

## Genome-wide Manhattan plot

Use `ez_manhattan()` to visualize GWAS results across the entire genome. We use the example GWAS data from `qqman` package to demonstrate.

```{r}
library(qqman)
head(gwasResults)
```

```{r}
ez_manhattan(
  input = gwasResults,
  chr = "CHR",
  bp = "BP",
  p = "P",
  colors = c("grey", "skyblue")
)
```

You can add a genome-wide significance threshold line:

```{r}
ez_manhattan(
  input = gwasResults,
  chr = "CHR",
  bp = "BP",
  p = "P",
  colors = c("grey", "skyblue"),
  threshold_p = 5e-8,
  threshold_color = "red"
)
```

## Regional association plot (LocusZoom-style)

Use `ez_locusZoom()` to create regional association plots focused on a single locus. This is ideal for fine-mapping visualization and can be stacked with other genomic tracks.

```{r}
# Filter data to a specific region
region <- "chr3:200-500"
locusResults <- gwasResults |>
  dplyr::filter(CHR == 3 & BP > 200 & BP < 500)
```

### Basic regional plot

```{r}
ez_locusZoom(
  input = locusResults,
  region = region,
  snp = "SNP",
  size = 1
)
```

### With LD coloring

You can use the `r2` option to color the dots by linkage disequilibrium with a lead variant. This creates the classic LocusZoom color gradient. For demonstration purpose, here we use p-value to create a continuous color for `r2`.

```{r}
# Simulated r2 values (in practice, these come from LD calculations)
r2_values <- -log10(locusResults$P) / max(-log10(locusResults$P))

ez_locusZoom(
  input = locusResults,
  region = region,
  snp = "SNP",
  lead_snp = "rs3057",
  r2 = r2_values,
  size = 2
)
```
