---
title: "Sequence track"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(ezGenomeTracks)
library(ggplot2)
```

`ezGenomeTracks` provides `ez_sequence()` to plot nucleotide sequence from a reference genome. Each base is colored following [UCSC Genome Browser](https://genome.ucsc.edu/) conventions: **A** (green), **C** (blue), **G** (gold), **T** (red), **N** (grey).

## Setup

You need a BSgenome data package for your organism. For human hg38:

```{r eval=FALSE}
BiocManager::install("BSgenome.Hsapiens.UCSC.hg38")
```

```{r}
library(BSgenome.Hsapiens.UCSC.hg38)
genome <- BSgenome.Hsapiens.UCSC.hg38::Hsapiens
```

Define a short region to inspect (sequence tracks are most useful at base-pair resolution):

```{r}
region <- "chr2:231109000-231109030"
```

## Visual styles

`ez_sequence()` offers two visual styles controlled by the `style` argument.

### Text style (default)

The default `style = "text"` draws bold, UCSC-colored nucleotide letters with no background or border — a clean look that is readable at narrow regions and blends well with other tracks.

```{r, fig.height=0.5}
ez_sequence(genome, region = region)
```

### Tile style

`style = "tile"` draws colored background tiles with white letter labels on top, echoing the classic UCSC "dense" sequence display.

```{r, fig.height=0.5}
ez_sequence(genome, region = region, style = "tile")
```

## Automatic label suppression

When the region is wider than `max_label_bp` (default 200 bp), `show_labels = "auto"` (the default) suppresses individual letter labels automatically. This prevents overplotting while still rendering colored blocks for orientation.

```{r, fig.height=0.5}
# Wide region: labels hidden automatically, tiles drawn instead
ez_sequence(genome, region = "chr2:231109000-231109500", style = "tile")
```

You can override this behaviour by setting `show_labels` explicitly:

```{r, fig.height=0.5}
# Force labels off regardless of region width
ez_sequence(genome, region = region, show_labels = FALSE, style = "tile")
```

## Customizing colors

Use the `colors` argument to override individual nucleotide colors. Only the specified bases are changed; others retain UCSC defaults.

```{r, fig.height=0.5}
ez_sequence(
  genome,
  region = region,
  colors = c(A = "darkorchid", T = "deeppink")
)
```

Inspect the full default palette:

```{r}
ez_sequence_palette()
```

Replace the entire palette by supplying all five bases:

```{r, fig.height=0.5}
grayscale <- c(
  A = "#222222",
  C = "#555555",
  G = "#888888",
  T = "#BBBBBB",
  N = "#DDDDDD"
)
ez_sequence(genome, region = region, colors = grayscale)
```

## Adjusting label size and tile height

```{r, fig.height=0.5}
ez_sequence(
  genome,
  region = region,
  style = "tile",
  label_size = 4,
  tile_height = 1
)
```

## Stacking with other tracks

`ez_sequence()` returns a standard ggplot2 object and is fully compatible with `vstack_plot()`. A common pattern is to place a sequence track at the bottom of a stacked plot for base-level context.

```{r}
bw0 <- system.file(
  "extdata",
  "avg_chr2-231091223_231109786_231113600_0.bw",
  package = "ezGenomeTracks"
)
```

```{r, fig.height=2.5}
cov_track <- ez_coverage(bw0, region = region, y_axis_style = "simple")
seq_track <- ez_sequence(genome, region = region)

vstack_plot(cov_track, seq_track, heights = c(1))
```

The tile style can also be useful in a stacked context, particularly when the region is very narrow:

```{r, fig.height=2.5}
vstack_plot(
  ez_coverage(bw0, region = region, y_axis_style = "simple"),
  ez_sequence(genome, region = region, style = "tile"),
  heights = c(1)
)
```

## Low-level usage with `geom_sequence`

For full control, you can use `geom_sequence()` directly in a `ggplot()` call. The expected data frame needs `position`, `nucleotide`, and `fill` columns, which you can build with `ez_sequence_palette()` and a DNAString:

```{r}
library(BSgenome)
library(GenomicRanges)

region_gr <- parse_region(region)
seq_char <- as.character(getSeq(genome, region_gr)[[1]])
nucs <- strsplit(seq_char, "")[[1]]
palette <- ez_sequence_palette()

seq_df <- data.frame(
  position = seq(
    GenomicRanges::start(region_gr),
    by = 1L,
    length.out = length(nucs)
  ),
  nucleotide = nucs,
  fill = palette[nucs]
)

head(seq_df)
```

```{r, fig.height=0.5}
ggplot(seq_df, aes(x = position, label = nucleotide, fill = fill)) +
  geom_sequence(label_size = 3.5) +
  scale_fill_identity() +
  scale_x_genome_region(region) +
  ylim(0, 1) +
  ez_sequence_theme()
```
