---
title: "Feature/Peak Track"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 2,
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(ezGenomeTracks)
library(ggplot2)
```

Feature tracks visualize genomic intervals such as ChIP-seq peaks, enhancers, promoters, or any other region-based annotations. The `ez_feature()` function provides a simple interface for creating these visualizations from BED files or data frames.

# Basic Usage

## From a Data Frame

The simplest way to create a feature track is from a data frame with genomic coordinates.

```{r basic-df}
# Create example peak data
peaks_df <- data.frame(
  seqnames = "chr1",
  start = c(1000, 3500, 6000, 8500),
  end = c(2000, 4500, 7500, 9500),
  name = c("peak1", "peak2", "peak3", "peak4"),
  score = c(100, 250, 180, 320)
)

# Basic feature track
ez_feature(peaks_df, region = "chr1:1-10000")
```

## From a BED File

You can also load features directly from a BED file.

```{r from-bed}
# Load example peaks from package
bed_file <- system.file("extdata", "example_peaks.bed", package = "ezGenomeTracks")

ez_feature(bed_file, region = "chr1:1-10000")
```

# Customizing Appearance

## Colors and Transparency

Customize the fill color, border color, and transparency.

```{r custom-colors}
ez_feature(
  peaks_df,
  region = "chr1:1-10000",
  fill = "darkgreen",
  color = "black",
  alpha = 0.8
)
```

```{r custom-colors-2}
# Different color scheme
ez_feature(
  peaks_df,
  region = "chr1:1-10000",
  fill = "coral",
  color = "darkred",
  alpha = 0.6
)
```

## Adjusting Feature Height

Control the height of feature rectangles (0 to 1 scale).

```{r height}
# Taller features
ez_feature(
  peaks_df,
  region = "chr1:1-10000",
  height = 0.9,
  fill = "steelblue"
)
```

```{r height-thin}
# Thinner features (useful when stacking with other tracks)
ez_feature(
  peaks_df,
  region = "chr1:1-10000",
  height = 0.4,
  fill = "purple"
)
```

# Score-Based Coloring

When your features have associated scores (e.g., peak significance, enrichment), you can map scores to color intensity.

```{r score-coloring}
# Color by score - gradient from white to fill color
ez_feature(
  peaks_df,
  region = "chr1:1-10000",
  use_score = TRUE,
  fill = "red"
)
```

```{r score-coloring-blue}
# Different gradient
ez_feature(
  peaks_df,
  region = "chr1:1-10000",
  use_score = TRUE,
  fill = "darkblue"
)
```

# Stacking with Other Tracks

Feature tracks are commonly used alongside coverage and gene tracks to show peak locations in genomic context.
```{r stacking, fig.height=5}
# Load coverage data
bw_file <- system.file(
  "extdata",
  "avg_chr2-231091223_231109786_231113600_0.bw",
  package = "ezGenomeTracks"
)

region <- "chr2:231109000-231115000"

# Create tracks
coverage_track <- ez_coverage(
  bw_file,
  region = region,
  colors = "steelblue",
  y_axis_style = "simple"
)

# Create synthetic peaks for this region
peaks_chr2 <- data.frame(
  seqnames = "chr2",
  start = c(231109500, 231111000, 231113000),
  end = c(231110500, 231112000, 231114000),
  score = c(80, 150, 120)
)

peak_track <- ez_feature(
  peaks_chr2,
  region = region,
  use_score = TRUE,
  fill = "darkred"
)

# Stack vertically (heights is relative height of second track vs first)
vstack_plot(
  coverage_track,
  peak_track,
  region = region,
  heights = c(0.3)
)
```

# Advanced Usage

## Using geom_feature() Directly

For more control, use `geom_feature()` directly with ggplot2.

```{r geom-direct}
ggplot(peaks_df) +

  geom_feature(
    aes(fill = score),
    height = 0.7,
    color = "gray30"
  ) +
  scale_fill_gradient(low = "lightyellow", high = "red") +
  scale_x_genome_region("chr1:1-10000") +
  ylim(0, 1) +
  theme_minimal() +
 theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(x = "Chr1", fill = "Score")
```

## Custom Aesthetics

Map other data columns to aesthetics.

```{r custom-aes}
# Add a category column
peaks_df$category <- c("promoter", "enhancer", "promoter", "enhancer")

ggplot(peaks_df) +
 geom_feature(
    aes(fill = category),
    height = 0.7,
    alpha = 0.8
  ) +
  scale_fill_manual(values = c("promoter" = "blue", "enhancer" = "orange")) +
  scale_x_genome_region("chr1:1-10000") +
  ylim(0, 1) +
  theme_void() +
  theme(legend.position = "bottom")
```

# Tips and Best Practices

1. **Height adjustment**: When stacking multiple tracks, use smaller `height` values (0.3-0.5) for feature tracks to maintain visual balance.

2. **Score normalization**: If scores vary widely, consider normalizing them before plotting or the color gradient may not be informative.

3. **Region filtering**: For large BED files, the function automatically filters to the specified region, but providing pre-filtered data improves performance.

4. **Overlapping features**: Features that overlap in genomic coordinates will be drawn on top of each other. Consider using transparency (`alpha`) to visualize overlaps.
