---
title: "Link/Arc Track"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 3,
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(ezGenomeTracks)
library(ggplot2)
```

Link/arc tracks visualize pairwise genomic interactions such as chromatin loops, enhancer-promoter connections, or any paired-end data. The `ez_link()` function draws curved arcs connecting two genomic regions.

> **Note**: Due to a rendering limitation with bezier curves in knitr/R Markdown documents, the code examples in this vignette are shown as code only. Please run these examples interactively in R to see the visualizations.

# Basic Usage

## Creating Interaction Data

Interaction data requires paired coordinates: `start1/end1` for the first anchor and `start2/end2` for the second anchor.

```{r basic, eval=FALSE}
# Create example interaction data
interactions_df <- data.frame(
  start1 = c(1000, 2500, 5000),
  end1 = c(1500, 3000, 5500),
  start2 = c(4000, 6000, 8000),
  end2 = c(4500, 6500, 8500),
  score = c(10, 25, 15)
)

# Basic arc track
ez_link(interactions_df, region = "chr1:1-10000")
```

## From Package Example Data

```{r example-data}
# Load example interactions
data(example_interactions)
head(example_interactions)
```

```{r example-plot, eval=FALSE}
ez_link(example_interactions, region = "chr1:1-15000")
```

# Arc Direction

Control whether arcs curve upward or downward.

```{r direction-down, eval=FALSE}
# Arcs curving downward (default)
ez_link(
  interactions_df,
  region = "chr1:1-10000",
  direction = "down",
  colors = "steelblue"
)
```

```{r direction-up, eval=FALSE}
# Arcs curving upward
ez_link(
  interactions_df,
  region = "chr1:1-10000",
  direction = "up",
  colors = "coral"
)
```

# Customizing Appearance

## Colors and Line Properties

```{r custom-style, eval=FALSE}
ez_link(
  interactions_df,
  region = "chr1:1-10000",
  colors = "purple",
  size = 1.5,
  alpha = 0.8
)
```

## Arc Shape

Control the curvature and height of arcs.

```{r curvature, eval=FALSE}
# Higher curvature = more pronounced curves
ez_link(
  interactions_df,
  region = "chr1:1-10000",
  curvature = 0.8,
  height_factor = 0.25,
  colors = "darkgreen"
)
```

```{r flat-arcs, eval=FALSE}
# Flatter arcs
ez_link(
  interactions_df,
  region = "chr1:1-10000",
  curvature = 0.3,
  height_factor = 0.1,
  colors = "darkblue"
)
```

# Score-Based Coloring

Map interaction scores (e.g., contact frequency, significance) to color intensity.

```{r score-color, eval=FALSE}
ez_link(
  interactions_df,
  region = "chr1:1-10000",
  use_score = TRUE,
  colors = "red",
  size = 1.2
)
```

# Multiple Tracks

## Grouped Interactions

Visualize interactions from different conditions or cell types together.

```{r grouped, eval=FALSE}
# Add group information
interactions_grouped <- rbind(
  transform(interactions_df, cell_type = "K562"),
  transform(
    data.frame(
      start1 = c(1500, 3500),
      end1 = c(2000, 4000),
      start2 = c(5000, 7500),
      end2 = c(5500, 8000),
      score = c(20, 30)
    ),
    cell_type = "GM12878"
  )
)

ez_link(
  interactions_grouped,
  region = "chr1:1-10000",
  group_var = "cell_type",
  colors = c("blue", "orange"),
  alpha = 0.7
)
```

## Stacked Tracks (Faceted)

Display interactions from different samples as separate tracks.

```{r stacked, fig.height=4, eval=FALSE}
# Create named list for faceted display
interaction_list <- list(
  "Sample A" = interactions_df,
  "Sample B" = data.frame(
    start1 = c(2000, 4500),
    end1 = c(2500, 5000),
    start2 = c(6000, 8500),
    end2 = c(6500, 9000),
    score = c(18, 22)
  )
)

ez_link(
  interaction_list,
  region = "chr1:1-10000",
  colors = c("purple", "darkgreen"),
  facet_label_position = "left"
)
```

# Combining with Other Tracks

Arc tracks are most informative when combined with coverage and gene tracks to show interactions in genomic context.

```{r combined, fig.height=6, eval=FALSE}
region <- "chr1:1-15000"

# Coverage track (using example data)
data(example_signal)
cov_track <- ez_coverage(
  example_signal,
  region = region,
  colors = "steelblue",
  y_axis_style = "simple"
)

# Arc track
arc_track <- ez_link(
  example_interactions,
  region = region,
  direction = "down",
  colors = "coral",
  size = 1
)

# Gene track
data(example_genes)
gene_track <- ez_gene(example_genes, region = region)

# Stack all tracks
vstack_plot(
  cov_track,
  arc_track,
  gene_track,
  region = region,
  heights = c(0.75, 0.5)
)
```

# Advanced Usage

## Using geom_link() Directly

For full customization, use `geom_link()` with ggplot2.

```{r geom-direct, eval=FALSE}
ggplot(interactions_df) +
  geom_link(
    aes(
      start1 = start1, end1 = end1,
      start2 = start2, end2 = end2,
      color = score
    ),
    direction = "down",
    height_factor = 0.2,
    linewidth = 1.2
  ) +
  scale_color_viridis_c(option = "plasma") +
  scale_x_genome_region("chr1:1-10000") +
  theme_minimal() +
  theme(
    axis.text.y = element_blank(),
    axis.ticks.y = element_blank(),
    panel.grid = element_blank()
  ) +
  labs(x = "Chr1", color = "Score")
```

# Tips and Best Practices

1. **Arc direction**: Use `direction = "down"` when placing arcs below coverage tracks, `"up"` when above.

2. **Height factor**: Adjust `height_factor` based on the genomic span of your interactions. Longer-range interactions may need smaller values to prevent arcs from becoming too tall.

3. **Score scaling**: For interaction scores with large ranges, consider log-transforming before plotting.

4. **Overlapping arcs**: Use transparency (`alpha`) when many arcs overlap to see density patterns.

5. **BEDPE format**: When loading from files, ensure your BEDPE file has columns for both anchor coordinates (chr1, start1, end1, chr2, start2, end2).

6. **Interactive use**: Due to rendering limitations with bezier curves in knitr documents, arc tracks are best viewed in interactive R sessions or saved directly to files using `ggsave()`.
