---
title: "Hi-C Track"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 10,
  fig.height = 5,
  message = FALSE,
  warning = FALSE
)
```

```{r setup}
library(ezGenomeTracks)
library(ggplot2)
```

Hi-C tracks visualize chromatin contact frequencies from chromosome conformation capture experiments. The `ez_hic()` function supports both triangle (rotated) and square heatmap views, with multiple color palettes and scale transformations.

# Basic Usage

## From a Matrix

The most common input is a symmetric contact matrix.

```{r basic-matrix}
# Create example symmetric matrix
set.seed(42)
n <- 50
mat <- matrix(0, n, n)
for (i in 1:n) {
  for (j in i:n) {
    # Simulate distance-dependent decay
    distance <- abs(j - i)
    value <- 100 * exp(-distance / 10) + rnorm(1, 0, 5)
    mat[i, j] <- max(0, value)
    mat[j, i] <- mat[i, j]
  }
}
# Add some TAD-like structure
mat[10:20, 10:20] <- mat[10:20, 10:20] + 30
mat[30:45, 30:45] <- mat[30:45, 30:45] + 25

# Plot as triangle (default)
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000
)
```

## From a Sparse Data Frame

Hi-C data can also be provided as a sparse data frame with position pairs and scores.

```{r sparse-df}
# Create sparse format
data(example_hic)
head(example_hic)
```

```{r sparse-plot}
ez_hic(
  example_hic,
  region = "chr1:1-500000",
  resolution = 10000
)
```

# Visualization Styles

## Triangle View (Default)

The triangle view rotates the matrix 45 degrees, with the diagonal at the bottom. This is the standard genome browser representation.

```{r triangle}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  style = "triangle"
)
```

## Square Heatmap View

The square view shows the full symmetric matrix, useful for detailed inspection.

```{r square}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  style = "square"
)
```

# Color Palettes

ezGenomeTracks provides several color palettes optimized for Hi-C visualization.

## Cooler (Red) - Default

```{r palette-cooler}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  palette = "cooler"
)
```

## Yellow-Green-Blue

```{r palette-ylgnbu}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  palette = "ylgnbu"
)
```

## Viridis

```{r palette-viridis}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  palette = "viridis"
)
```

## Blue-White-Red (Diverging)

Useful for differential Hi-C or log-ratio comparisons.

```{r palette-bwr}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  palette = "bwr"
)
```

# Scale Transformations

Hi-C data often spans several orders of magnitude. Use transformations to improve visualization.
## Log10 Transform

```{r transform-log10}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  transform = "log10"
)
```

## Square Root Transform

```{r transform-sqrt}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  transform = "sqrt"
)
```

# Filtering by Distance

For triangle views, you can limit the maximum interaction distance shown.

```{r max-distance}
# Only show interactions within 300kb
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  max_distance = 300000
)
```

# Setting Color Scale Limits

Control the color scale range to highlight specific value ranges or ensure consistency across samples.

```{r limits}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  transform = "log10",
  limits = c(0.5, 2)  # log10 scale limits
)
```

# Combining with Other Tracks

Hi-C tracks are typically displayed with coverage and gene tracks to interpret chromatin structure in genomic context.

```{r combined, fig.height=8}
region <- "chr1:1000000-1500000"
resolution <- 10000

# Hi-C track
hic_track <- ez_hic(
  mat,
  region = region,
  resolution = resolution,
  style = "triangle",
  max_distance = 300000,
  palette = "cooler"
)

# Simulated signal track
signal_df <- data.frame(
  seqnames = "chr1",
  start = seq(1000000, 1490000, by = 1000),
  end = seq(1001000, 1491000, by = 1000),
  score = abs(sin(seq(0, 10*pi, length.out = 491)) * 50 + rnorm(491, 20, 5))
)

signal_track <- ez_coverage(
  signal_df,
  region = region,
  colors = "steelblue",
  y_axis_style = "simple"
)

# Stack tracks
vstack_plot(
  hic_track,
  signal_track,
  region = region,
  heights = c(0.33)  # signal track is 1/3 height of Hi-C track
)
```

# Advanced Usage

## Rasterization for Large Matrices

For regions with many bins, rasterization improves performance and reduces file size.

```{r rasterize, eval=FALSE}
ez_hic(
  large_matrix,
  region = "chr1:1-10000000",
  resolution = 10000,
  rasterize = TRUE,
  rasterize_dpi = 300
)
```

## Using geom_hic_triangle() Directly

For full customization, use the geom directly with ggplot2.

```{r geom-direct}
# Convert matrix to data frame format
hic_df <- data.frame(
  pos1 = rep(seq(1000000, 1490000, by = 10000), each = 50),
  pos2 = rep(seq(1000000, 1490000, by = 10000), times = 50),
  score = as.vector(mat)
)
hic_df <- hic_df[hic_df$pos2 >= hic_df$pos1, ]  # Upper triangle only

ggplot(hic_df, aes(x = pos1, y = pos2, fill = score)) +
  geom_hic_triangle(resolution = 10000, max_distance = 300000) +
  scale_fill_hic(palette = "cooler", trans = "log10") +
  scale_x_genome_region("chr1:1000000-1500000") +
  coord_cartesian(clip = "off") +
  theme_minimal() +
  labs(x = "Chr1", fill = "Contacts")
```

## Hiding the Diagonal

Remove self-interactions (diagonal) from the plot.

```{r no-diagonal}
ez_hic(
  mat,
  region = "chr1:1000000-1500000",
  resolution = 10000,
  show_diagonal = FALSE
)
```

# Tips and Best Practices

1. **Resolution**: Always specify the correct resolution (bin size) of your Hi-C data. Common resolutions are 5kb, 10kb, 25kb, 50kb, and 100kb.

2. **Transformation**: For raw contact counts, `transform = "log10"` usually works best. For balanced/normalized data, `"identity"` or `"sqrt"` may be appropriate.

3. **Memory**: Large Hi-C matrices consuming memory. Use `max_distance` to limit data or downsample for large regions.

4. **Color limits**: Set consistent `limits` when comparing multiple samples to ensure colors are comparable.

5. **File formats**: For .hic or .cool files, use external tools (like `strawr` or `cooler`) to extract the region of interest before plotting.

6. **TAD visualization**: TAD boundaries often appear as corners of contact domains. Use lower resolutions (25-50kb) to see TAD structure clearly.
